---
author: "Hector Roux de Bézieux"
date: '`r format(Sys.time(), "%d %B , %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
title: ARI issues
---

```{r packages, include=F}
library(knitr)
opts_chunk$set(
  fig.pos = "!h", out.extra = "", warning = F, message = F,
  fig.width = 5, fig.align = "center"
)
libs <- c("here", "tidyverse", "cowplot", "clusterExperiment", "mclust")
suppressMessages(
  suppressWarnings(sapply(libs, require, character.only = TRUE))
)
rm(libs)
```

As mentionned by Elizabeth at the Brainstat Meeting on 03/11, ARI might not be the best metric.

One issue I see, especially when using it as a final metric, is that ARI is symmetric. Hence, we have the following situation.

```{r}
x <- rep(1:3, each = 10)
y <- c(rep(1, 20), rep(3, 10))

plotClusters(data.frame(x = x, y = y) %>% as.matrix())
```

The two cluster labels agree perfectly, but x gives a better resolution (i.e split a cluster in two).

In that case, the ARI is  `r round(adjustedRandIndex(x, y), 2)`.


This is what we have here.
