---
author: "Hector Roux de Bézieux"
date: '`r format(Sys.time(), "%d %B , %Y")`'
title: "Installing Monocle"
output: pdf_document
---

# Installing Monocle 2
## First Try

No module loaded.
Just open __R__ and do _BiocManager::install("monocle")_.

This fails for packages __densityClust__ and __DDRTree__.

## Second try

First do _module load gcc_ then try again. This works for __densityClust__ but not __DDRTree__.

## Third try 

I realized that this gives the same error as when trying to install _clusterExperiment_. 

```{r, eval = FALSE}
* installing *source* package ‘DDRTree’ ...
** package ‘DDRTree’ successfully unpacked and MD5 sums checked
** libs
icpc -std=gnu++11 -I"/opt/packages/R/3.5.1-mkl/lib64/R/include" 
-DNDEBUG  -I"/pylon5/ib5phhp/diyadas/rpack/3.5/Rcpp/include" 
-I"/pylon5/ib5phhp/hectorrb/rpack/RcppEigen/include" 
-I"/pylon5/ib5phhp/diyadas/rpack/3.5/BH/include" -I/usr/local/include   -fpic  -O3 
-ipo -openmp -xHost -c DDRTree.cpp -o DDRTree.o
icpc: command line error: option '-openmp' is not supported.
Please use the replacement option '-qopenmp'
make: *** [DDRTree.o] Error 1
ERROR: compilation failed for package ‘DDRTree’
* removing ‘/pylon5/ib5phhp/hectorrb/rpack/DDRTree’
```

I will try to use the same trick.

# Installing other packages
## Python packages


```{python, eval = FALSE}
pip install --user virtualenv
```

```{r, eval = FALSE}
install.packages("reticulate")
library(reticulate)
py_install('umap-learn')
```

```{python, eval = FALSE}
pip install --user louvain
```

Installing Louvain fails as well...